### 2015-07-31 doc/thesis/glossary.tex
Minor fixes to thesis text

\item[depth subtyping] An operation that allows variance in the type of
record fields.

@@

\item[type environment] An environment used by Typed Lua to assign variable names to first-level types.

@@

\item[variance] A way to define the subtyping order between the components
of a type constructor.

\item[width subtyping] An operation that allows adding fields to a record.

>>>>

\item[depth subtyping] An operation over records that allows variance in the type of
record fields.

@@

\item[type environment] An environment that binds variable names to types.

@@

\item[variance] A property that defines the subtyping order between the components
of a type constructor.

\item[width subtyping] An operation over records that allows the subtype to include
fields that do not exist in the supertype.

### 2015-07-31 doc/thesis/system.tex
Minor fixes to thesis text

In Section \ref{sec:refinement} we will show that table refinement can
change the type environment while typing an expression or a left-hand value.
Thus, the rules that type check lists of expressions and lists of
left-hand values use a partial auxiliary function \emph{merge} to collect
all environment changes in a new environment $\env_{f}$, if there are no conflicts.
In Section \ref{sec:refinement} we will also show that we can only change the
environment to add new table fields in a table type, and we cannot change the
type of a variable or a table field which is already present in a table type.

@@

Note that the coercion rules only allows changing the type
of a variable if the new type is a supertype of the previous type,
and the resulting type is always \emph{fixed} or \emph{closed}
to prevent the creation of unsafe aliases.
If the coercion is to a \emph{closed} table type the type of the
table changes to an \emph{open} table type with the same shape,
but if the coercion is to a \emph{fixed} table type the table
has to assume the same type.

@@

This rule uses the auxiliary function \emph{infer} to get the
most general types of each variable that should be introduced in
the type environment for type checking $s$.
After type checking the statement $s$, rule \textsc{T-LOCAL2} produces a
new type environment $\env_{3}$ without the variables that it introduced
before type checking $s$.
We can define \emph{infer} as follows:
\begin{align*}
infer(T_{1} \times ... \times T_{n}{*}, i) & = \left\{
\begin{array}{ll}
general(T_{i}) & \text{if $i < n$}\\
general(nil(T_{n})) & \text{if $i >= n$}
\end{array} \right.
\end{align*}

@@

The rule \textsc{T-IF5} shows the case where our type system
discriminates a projection type based on the tag \texttt{nil}.
It uses the auxiliary functions \emph{fopt} and \emph{fipt}
to filter a projection $x$, affecting all variables that bind to the same projection.
More precisely, the former function filters out the tuples that contain a type $\Nil$ in the $i$-th component,
while the latter function filters out the tuples that do not contain $\Nil$ in the $i$-th component.
We define \textsc{T-IF5} as follows:
\[
\begin{array}{c}
\mylabel{T-IF5}\\
\dfrac{\begin{array}{c}
       \env_{1}(id) = \pi_{i}^{x} \\
       S_{t} = fopt(\penv(x), \Nil, i) \;\;\;
       S_{e} = fipt(\penv(x), \Nil, i) \\
       \env_{1}, \penv[x \mapsto S_{t}] \vdash s_{1}, \env_{2}\\
       \env_{1}, \penv[x \mapsto S_{e}] \vdash s_{2}, \env_{3}\\
       \env_{4} = join(\env_{2}, \env_{3})
      \end{array}}
      {\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{4}}
\end{array}
\]

As an example, our previous example type checks through rule \textsc{T-IF5},
because it uses the information provided by the projection type $\pi_{1}^{x}$,
which is the type of $q$, to make the rule \textsc{T-IF5} use the function call
\[
fopt((\Integer \times \Integer \times \Nil{*}) \sqcup (\Nil \times \String \times \Nil{*}), \Nil, 1)
\]
to discriminate the projection $x$ to the single tuple
$\Integer \times \Integer \times \Nil{*}$ inside the $\mathbf{if}$ branch,
and the function call
\[
fipt((\Integer \times \Integer \times \Nil{*}) \sqcup (\Nil \times \String \times \Nil{*}), \Nil, 1)
\]
to discriminate the projection $x$ to the single tuple
$\Nil \times \String \times \Nil{*}$ inside the $\mathbf{else}$ branch.
Thus, reading $q$ and $r$ projects $\pi_{1}^{x}$ to $\Integer$ and
$\pi_{2}^{x}$ to $\Integer$ inside the $\mathbf{if}$ branch,
but it projects $\pi_{1}^{x}$ to $\Nil$ and $\pi_{2}^{x}$ to $\String$
inside the $\mathbf{else}$ branch.
Outside the condition, $q$ and $r$ use the original projection, that is,
they project to $\Integer \cup \Nil$ and $\Integer \cup \String$, respectively.

>>>>

Later, in this section we will show that table refinement can
change the type environment while typing an expression or a left-hand value.
Thus, the rules that type check lists of expressions and lists of
left-hand values use a partial auxiliary function \emph{merge} to collect
all environment changes in a new environment $\env_{f}$, if there are no conflicts.
We will also show that we can only change the
environment to add new table fields in a table type, and we cannot change the
type of a variable or a table field which is already present in a table type.

@@

Note that the coercion rules only allow changing the type
of a variable if the new type is a supertype of the previous type,
and the resulting type is always \emph{fixed} or \emph{closed}
to prevent the creation of unsafe aliases.
If the coercion is to a \emph{closed} table type the type of the
table changes to an \emph{open} table type with the same shape,
but if the coercion is to a \emph{fixed} table type the table
has to assume the same type.

@@

This rule uses the auxiliary function \emph{infer} to get the
most general types of each variable that should be introduced in
the type environment for type checking $s$.
After type checking the statement $s$, rule \textsc{T-LOCAL2} produces a
new type environment $\env_{4}$ without the variables that it introduced
before type checking $s$.
We can define \emph{infer} as follows:
\begin{align*}
infer(T_{1} \times ... \times T_{n}{*}, i) & = \left\{
\begin{array}{ll}
general(T_{i}) & \text{if $i < n$}\\
general(nil(T_{n})) & \text{if $i >= n$}
\end{array} \right.
\end{align*}

@@

The rule \textsc{T-IF5} shows the case where our type system
discriminates a projection type based on the tag \texttt{nil}.
It uses the auxiliary functions \emph{fopt} and \emph{fipt}
to filter a projection $x$, affecting all variables that bind to the same projection.
More precisely, the former function filters out the tuples that contain a type $F$ in the $i$-th component,
while the latter function filters out the tuples that do not contain $F$ in the $i$-th component.
We define \textsc{T-IF5} as follows:
\[
\begin{array}{c}
\mylabel{T-IF5}\\
\dfrac{\begin{array}{c}
       \env_{1}(id) = \pi_{i}^{x} \\
       S_{t} = fopt(\penv(x), \Nil, i) \;\;\;
       S_{e} = fipt(\penv(x), \Nil, i) \\
       \env_{1}, \penv[x \mapsto S_{t}] \vdash s_{1}, \env_{2}\\
       \env_{1}, \penv[x \mapsto S_{e}] \vdash s_{2}, \env_{3}\\
       \env_{4} = join(\env_{2}, \env_{3})
      \end{array}}
      {\env_{1}, \penv \vdash \mathbf{if} \; id \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}, \env_{4}}
\end{array}
\]

Our previous example type checks through rule \textsc{T-IF5},
because it uses the information provided by the projection type $\pi_{1}^{x}$,
which is the type of $q$, to make the rule \textsc{T-IF5} use the function call
\[
fopt((\Integer \times \Integer \times \Nil{*}) \sqcup (\Nil \times \String \times \Nil{*}), \Nil, 1)
\]
to discriminate the projection $x$ to the single tuple
$\Integer \times \Integer \times \Nil{*}$ inside the $\mathbf{if}$ branch,
and the function call
\[
fipt((\Integer \times \Integer \times \Nil{*}) \sqcup (\Nil \times \String \times \Nil{*}), \Nil, 1)
\]
to discriminate the projection $x$ to the single tuple
$\Nil \times \String \times \Nil{*}$ inside the $\mathbf{else}$ branch.
Thus, reading $q$ and $r$ projects $\pi_{1}^{x}$ to $\Integer$ and
$\pi_{2}^{x}$ to $\Integer$ inside the $\mathbf{if}$ branch,
but it projects $\pi_{1}^{x}$ to $\Nil$ and $\pi_{2}^{x}$ to $\String$
inside the $\mathbf{else}$ branch.
Outside the condition, $q$ and $r$ use the original projection, that is,
they project to $\Integer \cup \Nil$ and $\Integer \cup \String$, respectively.
